---
title: "Measures of Fit activity"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

## Setup
```{r eval = F}
## run this just once to get the packages
install.packages(c("corrplot", "manipulate"))

```

```{r setup}
library(lavaan)
library(manipulate)
library(corrplot)

tr = function (m) { #calc the trace of a matrix
  sum(diag(m))
}

discrepancy = function(obs, mod) {
 ## Calculate the ML discrepancy statistic, "F", between the observed and
 ## model-implied covariance matrices
  log(det(obs)) + tr(mod %*% solve(obs)) - log(det(mod)) - length(diag(obs))
}


## Sample original covariance matrix
source_cov = cov(HolzingerSwineford1939[, c("x1", "x4", "x7")])

srmr = function(obs, mod) {
  ## Calculate SRMR; based on https://stats.stackexchange.com/questions/235872/what-is-the-formula-for-standardized-root-mean-residual-srmr-in-the-context-of
  lobs <-  cov2cor(obs)[!lower.tri(obs)]
  limp <-  cov2cor(mod)[!lower.tri(mod)]
  sqrt(mean((limp - lobs)^2))
}

MAD_cor = function(obs, mod) {
  ## mean absolute correlation difference
  lobs <-  cov2cor(obs)[!lower.tri(obs)]
  limp <-  cov2cor(mod)[!lower.tri(mod)]
  mean(abs(limp - lobs))
}


## Baseline (independence)  covariance matrix -----
## diag() on a matrix returns a vector of the diagonal entries;
## diag() on a vector creates a matrix that has that diagonal
## Therefore diag(diag(matrix)) just sets everything besides the diagonal to 0
baseline_cov = diag(diag(source_cov))
baseline_F = discrepancy(source_cov, baseline_cov)
baseline_df = choose(nrow(source_cov), 2) # number of elements off diagonal

```


## Manipulate the chi-square

```{r}


manipulate({
  covmat = getCov(c(0, xy, 0, xz, yz, 0)) # get correlation differences
  cor_mod = cov2cor(source_cov) - covmat # calculate model-implied correlation
  cov_mod = cor2cov(cor_mod, sqrt(diag(source_cov))) # model-implied covariance
  calc_F = discrepancy(source_cov, cov_mod) # model discrepancy
  calc_chisq = n * calc_F # model chi-square
  pvalue = pchisq(calc_chisq, df, lower.tail = FALSE) # chi-square p-value
  
  ## Plot local fit and chi-square distribution
  par(mfrow = c(1, 2))
  corrplot(covmat, 
           addCoef.col = "black",
           title = paste0("Chisq=", round(calc_chisq, 1), " df=", df, " p=", round(pvalue, 3)),
           mar=c(0,0,1,0))
  par(mar = c(5, 2, 10, 2))
  curve(dchisq(x, df), from = 0, to = max(calc_chisq, df*4), main = "Chi-square distribution")
  abline(v = calc_chisq)
},
xy = slider(-.3, .3, 0),
xz = slider(-.3, .3, 0),
yz = slider(-.3, .3, 0),
df = slider(0, 5, 2),
n = slider(100, 1500, 200))

```

## With other global fit measures
Conceptually, SRMR is an average correlation difference; I've included a made-up statistic, the MAD cor, which is the mean absolute differences in the correlation matrices for comparison.

Note how CFI changes as we change parameters, relative to the baseline model, but doesn't change that much based on sample size. "Deltas" are the difference between the chi-square statistic and the degrees of freedom.

```{r}
manipulate({
  covmat = getCov(c(0, xy, 0, xz, yz, 0)) # get correlation differences
  cor_mod = cov2cor(source_cov) - covmat # calculate model-implied correlation
  cov_mod = cor2cov(cor_mod, sqrt(diag(source_cov))) # model-implied covariance
  calc_F = discrepancy(source_cov, cov_mod) # model discrepancy
  calc_chisq = n * calc_F # model chi-square
  pvalue = pchisq(calc_chisq, df, lower.tail = FALSE) # chi-square p-value
  baseline_chisq = n * baseline_F
  baseline_pvalue = pchisq(baseline_chisq, df, lower.tail = FALSE) #baseline p-value
  
  model_delta = max(0, calc_chisq - df)
  baseline_delta = baseline_chisq - baseline_df
  
  CFI = (baseline_delta - model_delta)/baseline_delta
  SRMR = srmr(source_cov, cov_mod)
  MAD = MAD_cor(source_cov, cov_mod)
  RMSEA = sqrt(model_delta/(df * (n - 1)))
  
  ## Plot local fit and chi-square distribution
  par(mfrow = c(1, 2))
  corrplot(covmat, 
           addCoef.col = "black",
           mar=c(0,0,1,0))
  title(sprintf("Chisq=%s, df=%s, p=%s\nBaseline Chisq=%s, df=%s, p=%s\nSRMR=%s, MAD cor=%s\nRMSEA=%s", 
                           round(calc_chisq, 1), 
                           df, 
                           round(pvalue, 3),
                           round(baseline_chisq, 1),
                           baseline_df,
                           round(baseline_pvalue, 3),
                round(SRMR, 3),
                round(MAD, 3),
                round(RMSEA, 3)))
  par(mar = c(5, 2, 10, 2))
  curve(dchisq(x, df), from = 0, to = max(calc_chisq, df*4), 
        main = sprintf("Chi-square distribution\nDeltas: Baseline=%s, Model=%s\nCFI = %s", 
                       round(baseline_delta, 1), round(model_delta, 1),
                       round(CFI,3)))
  abline(v = calc_chisq)
},
xy = slider(-.3, .3, 0),
xz = slider(-.3, .3, 0),
yz = slider(-.3, .3, 0),
df = slider(1, 2, 1),
n = slider(100, 1500, 200))

```

